Notes on Perturbational Quantities

### 2008-10-24 docs/pert_notes/pert.tex
addressed marginpar about predict_rho not being an option anymore

%-----------------------------------------------------------------------------
% density
%-----------------------------------------------------------------------------
\section{Density Evolution}
The full density evolution equation is
\begin{equation}
\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub) = -\Ub\cdot\nabla\rho 
- \rho\nabla\cdot\Ub. \label{rho equation}
\end{equation}

\subsection{Predicting $\rho'$ ({\tt predict\_rho = F} - default)}
The base state density evolution equation is \MarginPar{we don't have {\tt predict\_rho} anymore--is this still right?}
\begin{equation}
\frac{\partial\rho_0}{\partial t} = -\nabla\cdot(\rho_0 w_0 \eb_r) = 
-w_0\frac{\partial\rho_0}{\partial r} 
\underbrace{-\rho_0\frac{\partial w_0}{\partial r}}_{"\rho_0 ~ \text{force}"}.
\label{rho0 equation}
\end{equation}
Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange
terms, noting that $\Ub = \Ubt + w_o\eb_r$, to obtain the
perturbational density equation,
\begin{equation}
\frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub 
- \nabla\cdot(\rho_0\Ubt)}_{"\rho' ~ \text{force}"}. 
\label{rhoprime equation}
\end{equation}

\subsubsection{Predicting $\rho'$ at edges}
We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to
edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the
underbraced term in (\ref{rhoprime equation}) as the forcing.

\subsubsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
There are two ways to predict $\rho_0$ at edges.
\begin{enumerate}
\item We call {\tt make\_edge\_state\_1d} using the underbraced term
in (\ref{rho0 equation}) as the forcing.  This gives us
$\rho_0^{n+\half,{\rm pred}}$.  This term is used to advect $\rho_0$
in {\tt advect\_base} and therefore must also be used to compute
$\eta_\rho$, which will be used in {\tt correct\_base}.
\item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n +
\rho_0^{(2)})/2$.  We compute $\rho_0^{(2)}$ from $\rho_0^n$ using
{\tt advect\_base}, which advances equation (\ref{rho0 equation})
through $\Delta t$ in time.  The $(2)$ in the superscript indicates
that we have not called {\tt correct\_base} yet, which computes
$\rho_0^{n+1}$ from $\rho_0^{(2)}$.  We use $\rho_0^{(2)}$ rather than
$\rho_0^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be
consistent with the fact that we do not include an $\eta_\rho$
contribution to compute $\rho'$ at edges.  $\rho_0^{n+\half,{\rm
avg}}$ is used to construct $\rho$ at edges from $\rho'$ at edges, and
this $\rho$ at edges is used to compute fluxes for $\rho X_k$.
\end{enumerate}

\subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$
to $\rho'$ since $\rho_0^{n+\half,{\rm avg}}$ is a cell-centered
quantity.  For the radial edges, we interpolate to obtain
$\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to
$\rho'$.

\subsubsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}
The evolution equation for $\rho X_k$, ignoring the reaction terms
that were already accounted for in {\tt react\_state}, and the
associated discretization is
\begin{equation}
\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
-\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}} 
+ \rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.
\end{equation}

>>>>

%-----------------------------------------------------------------------------
% density
%-----------------------------------------------------------------------------
\section{Density Evolution}
The full density evolution equation is
\begin{equation}
\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub) = -\Ub\cdot\nabla\rho 
- \rho\nabla\cdot\Ub. \label{rho equation}
\end{equation}
We predict both $\rho'$ and $\rho_0$ to edges and later reconstruct $\rho$.
The base state density evolution equation is
\begin{equation}
\frac{\partial\rho_0}{\partial t} = -\nabla\cdot(\rho_0 w_0 \eb_r) = 
-w_0\frac{\partial\rho_0}{\partial r} 
\underbrace{-\rho_0\frac{\partial w_0}{\partial r}}_{"\rho_0 ~ \text{force}"}.
\label{rho0 equation}
\end{equation}
Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange
terms, noting that $\Ub = \Ubt + w_o\eb_r$, to obtain the
perturbational density equation,
\begin{equation}
\frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub 
- \nabla\cdot(\rho_0\Ubt)}_{"\rho' ~ \text{force}"}. 
\label{rhoprime equation}
\end{equation}

\subsection{Predicting $\rho'$ at edges}
We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to
edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the
underbraced term in (\ref{rhoprime equation}) as the forcing.

\subsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
There are two ways to predict $\rho_0$ at edges.
\begin{enumerate}
\item We call {\tt make\_edge\_state\_1d} using the underbraced term
in (\ref{rho0 equation}) as the forcing.  This gives us
$\rho_0^{n+\half,{\rm pred}}$.  This term is used to advect $\rho_0$
in {\tt advect\_base} and therefore must also be used to compute
$\eta_\rho$, which will be used in {\tt correct\_base}.
\item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n +
\rho_0^{(2)})/2$.  We compute $\rho_0^{(2)}$ from $\rho_0^n$ using
{\tt advect\_base}, which advances equation (\ref{rho0 equation})
through $\Delta t$ in time.  The $(2)$ in the superscript indicates
that we have not called {\tt correct\_base} yet, which computes
$\rho_0^{n+1}$ from $\rho_0^{(2)}$.  We use $\rho_0^{(2)}$ rather than
$\rho_0^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be
consistent with the fact that we do not include an $\eta_\rho$
contribution to compute $\rho'$ at edges.  $\rho_0^{n+\half,{\rm
avg}}$ is used to construct $\rho$ at edges from $\rho'$ at edges, and
this $\rho$ at edges is used to compute fluxes for $\rho X_k$.
\end{enumerate}

\subsection{Computing $\rho$ at edges}\label{Computing rho at edges}
For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$
to $\rho'$ since $\rho_0^{n+\half,{\rm avg}}$ is a cell-centered
quantity.  For the radial edges, we interpolate to obtain
$\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to
$\rho'$.

\subsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}
The evolution equation for $\rho X_k$, ignoring the reaction terms
that were already accounted for in {\tt react\_state}, and the
associated discretization is
\begin{equation}
\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
-\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}} 
+ \rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.
\end{equation}
